---
title: "progress-report-Alinayjk"
author: "Alina Kirkpatrick"
date: "2026-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Imports:
```{r}
library(rvest)
library(dplyr)
library(readr)
library(Lahman)

```

Scraping(making the scrapper):
```{r}
url <- "https://www.baseball-reference.com/awards/hof_2026.shtml"

bbwaa_scraper <- function(url, year) {
  site <- read_html(url)
  data <- site %>%
    html_element("#hof_BBWAA") %>%
    html_table()
  colnames(data) <- data[1, ]
  data <- data[-1, ]
  
  data <- data %>%
    select(Name, Votes, `%vote`) %>%
    mutate(
      yearID = year,
      Votes = as.numeric(Votes),
      percent_vote = parse_number(`%vote`),
      Name = gsub("[*#]", "", Name),
      inducted = ifelse(percent_vote >= 75, "Y", "N")
    )
  return(data)

}

```
Comment: I fixed messy headers form the BBWAA voting table, keept only Name, Votes and %vote as they are most relevant, converted Votes to numeric to ensure consistency and proper calculation, as well as %vote (where I used parse_number) and created yearID to help filter and inducted which checks the 75% or greater rule



Scraping 2026 data:
```{r}
url_2026 <- "https://www.baseball-reference.com/awards/hof_2026.shtml"

hof_2026 <- bbwaa_scraper(url_2026, 2026)

head(hof_2026)

summary(hof_2026)
```

lahman:
```{r}
lahman_2026 <- HallOfFame %>%
  filter(yearID == 2026)
```


comparing:
```{r}
hof_2026 %>%
  count(inducted)
#N = 25, Y = 2
lahman_2026 %>%
  count(votedBy, inducted)
#N = 7, Y = 1
setdiff(hof_2026$Name, lahman_2026$name)
#[1] "Carlos Beltrán"      "Andruw Jones"        "Chase Utley"        
# [4] "Andy Pettitte"       "Félix Hernández"     "Alex Rodriguez"     
# [7] "X-Manny Ramirez"     "Bobby Abreu"         "Jimmy Rollins"      
#[10] "Cole Hamels"         "Dustin Pedroia"      "Mark Buehrle"       
#[13] "Omar Vizquel"        "David Wright"        "Francisco Rodríguez"
#[16] "Torii Hunter"        "X-Ryan Braun"        "X-Edwin Encarnación"
#[19] "X-Shin-Soo Choo"     "X-Hunter Pence"      "X-Matt Kemp"        
#[22] "X-Rick Porcello"     "X-Nick Markakis"     "X-Alex Gordon"      
#[25] "X-Daniel Murphy"     "X-Howie Kendrick"    "X-Gio Gonzalez" 
```
Comment: the lahman data is very different, which is because the website shows the full ballot while Lahman doesn't. Lahman also uses playerID instead of names which makes comparrison dificult if you do not convert anything.


now save:
```{r}
write_csv(hof_2026, "hof_2026.csv")
```
